---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true

## Loading and preprocessing the data

Step One: Load in the data
```{r}
df <- read.csv("activity.csv")
```
Step Two: Check the head of the data
```{r}
head(df)

```
## What is mean total number of steps taken per day?
```{r}
library(ggplot2)

df.steps <- tapply(df$steps, df$date, Fun=sum, na.rm=True)

qplot(df.steps,xlab="Total steps per day",ylab="Freq using banwidth",binwidth= 1000)
```

The next step is to find the mean total number of steps per day.

```{r}      
mean(df.steps, na.rm=TRUE)
median(df.steps, na.rm=TRUE)
```
## What is the average daily activity pattern?
```{r}
df.averages <- aggregate(x=list(steps=df$steps), by=list(interval=df$interval), FUN=mean, na.rm=TRUE)
ggplot(data=df.averages, aes(x=interval, y=steps)) + geom_line() + xlab("Intervals set at 5 minutes") + ylab("Average of steps taken")
```
Further, on average for all days in the dataset df, the 5 minute intervals contains the following maximum number of steps:

```{r}
df.averages[which.max(df.averages$steps),]
```
## Imputing missing values
```{r}
df.missing <- is.na(df$steps)
table(df.missing)
nafiller <- function(steps, interval){
  filler <- NA
  if (!is.na(steps))
    filler <- c(steps)
  else
    filler <- (df.averages[df.averages$interval==interval, "steps"])
  return(filler)
}
myfill.df <- df
myfill.df$steps <- mapply(nafiller, myfill.df$steps, myfill.df$interval)
head(myfill.df)
```
The next thing we can do is utilize the histogram for visualization with the filled in data set.
```{r}
myts <- tapply(myfill.df$steps, myfill.df$date, FUN=sum)
qplot(myts, binwidth=1000, xlab="Total Number of Steps per Day",main="Total Number of Steps per Day After Imputation")
mean(myts)
median(myts)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
week.identify <- function(date){
 day <- weekdays(date)
 if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
   return("Weekday")
 else if (day %in% c("Saturday", "Sunday"))
   return("Weekend")
 else
   stop("Invalid Date")
}  
myfill.df$date <- as.Date(myfill.df$date)
myfill.df$day <- sapply(myfill.df$date, FUN=week.identify)
head(myfill.df$day)
avg <- aggregate(steps ~ interval + day, data=myfill.df, mean)
ggplot(avg, aes(interval, steps))+geom_line()+ facet_grid(day ~ .) + xlab("Intervals at 5 minutes") + ylab("# of Steps")
```

From the graph we see that weekday steps start out similar to the weekend steps. The difference is that more regular paterns occur in the weekend steps.